---
title: "p8105_hw2_dna2118"
author: "Dionna Attinson"
date: "9/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
## Question 1

### First we will clean the data Mr. Trash Wheel Data
```{r}
Healthy_Harbor = 
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "Mr. Trash Wheel") %>% 
  janitor::clean_names() %>% 
  select(-x15, -x16, -x17) %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = 
    as.integer(round(sports_balls, digit = 0)))
```
### Now, we will read and clean the 2017 precipitation data 
```{r}
Precip2017 = 
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2017 Precipitation",  
  range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(
    Year = 2017
  )
```
### Now, we will read and clean the 2018 precipitation data 
```{r}
Precip2018 = 
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2018 Precipitation",
  range="A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(
    Year = 2018
  )
```
### Now, we will combine the 2017 and 2018 precipitation datasets 
```{r}
Precip_Total = 
  full_join(Precip2017, Precip2018) %>% 
  janitor::clean_names() %>% 
  mutate(month=month.name[month])
```
### Summarizing the Data
The number of observations in the Healthy Harbor dataset is `r nrow(Healthy_Harbor)`. The number of observations in the joined 2017 and 2018 Precipitation dataset is `r nrow(Precip_Total)`. The total preciptation in 2018 was `r sum(pull(Precip2018,total))`. The median number of sports balls is `r median(pull(Healthy_Harbor, sports_balls))`. Some key variables that were present in both datasets include month and year. 