---
title: "p8105_hw2_dna2118"
author: "Dionna Attinson"
date: "9/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
## Question 1

### First we will clean the data Mr. Trash Wheel Data
```{r}
Healthy_Harbor = 
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "Mr. Trash Wheel") %>% 
  janitor::clean_names() %>% 
  select(-x15, -x16, -x17) %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = 
    as.integer(round(sports_balls, digit = 0)))
```
### Now, we will read and clean the 2017 precipitation data 
```{r}
Precip2017 = 
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2017 Precipitation",  
  range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(
    Year = 2017
  )
```
### Now, we will read and clean the 2018 precipitation data 
```{r}
Precip2018 = 
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2018 Precipitation",
  range="A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(
    Year = 2018
  )
```
### Now, we will combine the 2017 and 2018 precipitation datasets 
```{r}
Precip_Total = 
  full_join(Precip2017, Precip2018) %>% 
  janitor::clean_names() %>% 
  mutate(month=month.name[month])
```
### Summarizing the Data
The total number of observations in the Healthy Harbor dataset is `r nrow(Healthy_Harbor)`. The number of observations in the joined 2017 and 2018 precipitation data is `r nrow(Precip_Total)`. The total preciptation in 2018 was `r sum(pull(Precip2018,total))`. The median number of sports balls in a dumpster from 2017 `r median(pull(filter(Healthy_Harbor, year==2017), sports_balls))`. Some key variables that were present in both datasets include total, month and year. 

## Question 2
```{r}
pols_month = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"),
  convert = TRUE) %>% 
  mutate(month=month.name[month]) %>% 
  mutate(prez_dem=recode(prez_dem, "1" = "dem")) %>% 
  mutate(prez_gop=recode(prez_gop, "1" = "gop")) %>% 
  pivot_longer(c("prez_dem","prez_gop"),
    values_to = "president",
    values_drop_na = TRUE) %>% 
    select(-day)
```

```{r}
snp_data = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into= c("month","day","year"),
  convert = TRUE) %>% 
  select(year, month, -day, close) %>% 
  mutate(month=month.name[month])
```

```{r}
unemployment_data = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month", 
    values_to = "rate") %>% 
    mutate(month=match(month, month.abb)) %>% 
    mutate(month=month.name[month])  %>% 
    mutate(Year = as.integer(Year)) %>% 
    janitor::clean_names()
```

```{r}
P2_joined_data = 
  pols_month %>% 
  left_join(snp_data, 
    by=c("year", "month")) %>%   
  left_join(unemployment_data,
    by=c("year", "month"))
```
### Summarizing the data
The data pols_month contains 817 observations of 10 variables related to the number of national politicians who are democratic or republican at any given time. The file “unemployment” contains 816 observations of 4 variables that describe the percentage of unemployment in each month of the associated year.The file snp contains 787 observations of 3 variables related to Standard & Poor’s stock market index (S&P), often used as a representative measure of stock market as a whole. The key variables used are year and month.
